### 2021-11-30 Figures/Supp_figures.Rmd

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(reshape2)
library(EpiDISH)
library(tidyverse)
library(Hmisc)
library(gridExtra)
library(grid)

knitr::opts_chunk$set(fig.width = 20, fig.height = 30, cache = T)
```

### Silico1700

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
silico1700_facs <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Facs_Silico_1700.txt")
silico1700_deconv <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_Silico_1700.txt")

colnames(silico1700_deconv) <- colnames(silico1700_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(silico1700_deconv) <- colnames(silico1700_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(silico1700_deconv) <- colnames(silico1700_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(silico1700_deconv) <- colnames(silico1700_deconv) %>% str_replace_all("CBSX__LM22_", "CBSX.LM22")

silico1700_deconv <- silico1700_deconv %>% select(-colnames(silico1700_deconv)[grepl("XCELL", colnames(silico1700_deconv))])
silico1700_deconv <- silico1700_deconv %>% select(-colnames(silico1700_deconv)[grepl("MCP", colnames(silico1700_deconv))])

# CD4
cd4 <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("CD4", colnames(silico1700_deconv))])
# Regroup and remove different types of CD4 from LM22 signature
data$Epidish_CBSX_LM22_CD4 = data$Epidish_CBSX_LM22_T.cells.CD4.memory.activated + data$Epidish_CBSX_LM22_T.cells.CD4.memory.resting + data$Epidish_CBSX_LM22_T.cells.CD4.naive
data$DeconRNASeq_CBSX_LM22_CD4 = data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.naive
data$CBSX.LM22_CD4 = data$CBSX.LM22_T_cells_CD4_memory_activated + data$CBSX.LM22_T_cells_CD4_memory_resting + data$CBSX.LM22_T_cells_CD4_naive
data <- data %>% select(-c("CBSX.LM22_T_cells_CD4_memory_activated", "CBSX.LM22_T_cells_CD4_memory_resting", "CBSX.LM22_T_cells_CD4_naive", 
                           "Epidish_CBSX_LM22_T.cells.CD4.memory.activated" ,"Epidish_CBSX_LM22_T.cells.CD4.memory.resting" ,"Epidish_CBSX_LM22_T.cells.CD4.naive", 
                           "DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated" ,"DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting" , "DeconRNASeq_CBSX_LM22_T.cells.CD4.naive"))

data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$T.cells.CD4)

for (i in 1:(ncol(data) - 1)) {
  cd4[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD4.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd4, ncol = 3, top = textGrob("CD4", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# CD8
cd8 <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("CD8", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$T.cells.CD8)
for (i in 1:(ncol(data) - 1)) {
  cd8[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD8.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd8, ncol = 3, top = textGrob("CD8", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# Monocytes
mono <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("Mono", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$Monocytes)
for (i in 1:(ncol(data) - 1)) {
  mono[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_Mono.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = mono, ncol = 3, top = textGrob("Monocytes", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# M1
m1 <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("M1", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$Macrophages.M1)
for (i in 1:(ncol(data) - 1)) {
  m1[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(M1)|(_Macro.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = m1, ncol = 3, top = textGrob("Macrophages M1", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# M2
m2 <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("(_M2)|(\\.M2)", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$Macrophages.M2)
for (i in 1:(ncol(data) - 1)) {
  m2[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(M2)|(_Macro.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = m2, ncol = 3, top = textGrob("Macrophages M2", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# B
b <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("(B_cell)|(B.cell)|(B_lin)", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$B.cells)
for (i in 1:(ncol(data) - 1)) {
  b[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(B_cell.*)|(_B.cell.*)|(_B_lin.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = b, ncol = 3, top = textGrob("B cells", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# NK
nk <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("NK", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$NK.cells)
for (i in 1:(ncol(data) - 1)) {
  nk[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_NK.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = nk, ncol = 3, top = textGrob("Natural Killers", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# neutro
neutro <- list()
data <- silico1700_deconv %>% select(colnames(silico1700_deconv)[grepl("Neu", colnames(silico1700_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, silico1700_facs$Neutrophils)
for (i in 1:(ncol(data) -1)) {
  neutro[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_Neu.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = neutro, ncol = 3, top = textGrob("Neutrophils", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))
```

### GSE104171 - multiple myeloma

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

GSE104171_deconv <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_GSE104171.txt")
GSE104171_facs <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/FACS_GSE104171.txt")
GSE104171_facs[ 2:4] <- GSE104171_facs[ 2:4] / 100
colnames(GSE104171_deconv) <- colnames(GSE104171_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(GSE104171_deconv) <- colnames(GSE104171_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(GSE104171_deconv) <- colnames(GSE104171_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(GSE104171_deconv) <- colnames(GSE104171_deconv) %>% str_replace_all("CBSX__LM22_", "CBSX.LM22")

GSE104171_deconv <- GSE104171_deconv %>% select(-colnames(GSE104171_deconv)[grepl("XCELL", colnames(GSE104171_deconv))])
GSE104171_deconv <- GSE104171_deconv %>% select(-colnames(GSE104171_deconv)[grepl("MCP", colnames(GSE104171_deconv))])

# CD4
cd4 <- list()
data <- GSE104171_deconv %>% select(colnames(GSE104171_deconv)[grepl("CD4", colnames(GSE104171_deconv))])
# Regroup and remove different types of CD4 from LM22 signature
data$Epidish_CBSX_LM22_CD4 = data$Epidish_CBSX_LM22_T.cells.CD4.memory.activated + data$Epidish_CBSX_LM22_T.cells.CD4.memory.resting + data$Epidish_CBSX_LM22_T.cells.CD4.naive
data$DeconRNASeq_CBSX_LM22_CD4 = data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.naive
data$CBSX.LM22_CD4 = data$CBSX.LM22_T_cells_CD4_memory_activated + data$CBSX.LM22_T_cells_CD4_memory_resting + data$CBSX.LM22_T_cells_CD4_naive
data <- data %>% select(-c("CBSX.LM22_T_cells_CD4_memory_activated", "CBSX.LM22_T_cells_CD4_memory_resting", "CBSX.LM22_T_cells_CD4_naive", 
                           "Epidish_CBSX_LM22_T.cells.CD4.memory.activated" ,"Epidish_CBSX_LM22_T.cells.CD4.memory.resting" ,"Epidish_CBSX_LM22_T.cells.CD4.naive", 
                           "DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated" ,"DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting" , "DeconRNASeq_CBSX_LM22_T.cells.CD4.naive"))

data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE104171_facs$CD4)
for (i in 1:(ncol(data) - 1)) {
  cd4[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD4.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd4, ncol = 3, top = textGrob("CD4", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# CD8
cd8 <- list()
data <- GSE104171_deconv %>% select(colnames(GSE104171_deconv)[grepl("CD8", colnames(GSE104171_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE104171_facs$CD8)
for (i in 1:(ncol(data) - 1)) {
  cd8[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD8.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd8, ncol = 3, top = textGrob("CD8", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# NK
nk <- list()
data <- GSE104171_deconv %>% select(colnames(GSE104171_deconv)[grepl("NK", colnames(GSE104171_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE104171_facs$NK)
for (i in 1:(ncol(data) - 1)) {
  nk[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_NK.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = nk, ncol = 3, top = textGrob("Natural Killers", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

```

### Melanoma_GSE93722

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

GSE72056_facs <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Facs_Melanoma_GSE72056.txt")
GSE72056_deconv <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_Melanoma_GSE72056_no_malighant.txt")

colnames(GSE72056_deconv) <- colnames(GSE72056_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(GSE72056_deconv) <- colnames(GSE72056_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(GSE72056_deconv) <- colnames(GSE72056_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(GSE72056_deconv) <- colnames(GSE72056_deconv) %>% str_replace_all("CBSX__LM22_", "CBSX.LM22")

GSE72056_deconv <- GSE72056_deconv %>% select(-colnames(GSE72056_deconv)[grepl("XCELL", colnames(GSE72056_deconv))])
GSE72056_deconv <- GSE72056_deconv %>% select(-colnames(GSE72056_deconv)[grepl("MCP", colnames(GSE72056_deconv))])

# CD4
cd4 <- list()
data <- GSE72056_deconv %>% select(colnames(GSE72056_deconv)[grepl("CD4", colnames(GSE72056_deconv))])
# Regroup and remove different types of CD4 from LM22 signature
data$Epidish_CBSX_LM22_CD4 = data$Epidish_CBSX_LM22_T.cells.CD4.memory.activated + data$Epidish_CBSX_LM22_T.cells.CD4.memory.resting + data$Epidish_CBSX_LM22_T.cells.CD4.naive
data$DeconRNASeq_CBSX_LM22_CD4 = data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.naive
data$CBSX.LM22_CD4 = data$CBSX.LM22_T_cells_CD4_memory_activated + data$CBSX.LM22_T_cells_CD4_memory_resting + data$CBSX.LM22_T_cells_CD4_naive
data <- data %>% select(-c("CBSX.LM22_T_cells_CD4_memory_activated", "CBSX.LM22_T_cells_CD4_memory_resting", "CBSX.LM22_T_cells_CD4_naive", 
                           "Epidish_CBSX_LM22_T.cells.CD4.memory.activated" ,"Epidish_CBSX_LM22_T.cells.CD4.memory.resting" ,"Epidish_CBSX_LM22_T.cells.CD4.naive", 
                           "DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated" ,"DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting" , "DeconRNASeq_CBSX_LM22_T.cells.CD4.naive"))

data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE72056_facs$CD4)
for (i in 1:(ncol(data) - 1)) {
  cd4[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD4.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd4, ncol = 3, top = textGrob("CD4", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# CD8
cd8 <- list()
data <- GSE72056_deconv %>% select(colnames(GSE72056_deconv)[grepl("CD8", colnames(GSE72056_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE72056_facs$CD8)
for (i in 1:(ncol(data) - 1)) {
  cd8[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD8.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd8, ncol = 3, top = textGrob("CD8", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# Macrophages
macro <- list()
data <- GSE72056_deconv %>% select(colnames(GSE72056_deconv)[grepl("Macro", colnames(GSE72056_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE72056_facs$Macrophages)
for (i in 1:(ncol(data) - 1)) {
  macro[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_Macro.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = macro, ncol = 3, top = textGrob("Macrophages", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# B
b <- list()
data <- GSE72056_deconv %>% select(colnames(GSE72056_deconv)[grepl("(B_cell)|(B.cell)|(B_lin)", colnames(GSE72056_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE72056_facs$B)
for (i in 1:(ncol(data) - 1)) {
  b[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(B_cell.*)|(_B.cell.*)|(_B_lin.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = b, ncol = 3, top = textGrob("B cells", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# NK
nk <- list()
data <- GSE72056_deconv %>% select(colnames(GSE72056_deconv)[grepl("NK", colnames(GSE72056_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE72056_facs$NK)
for (i in 1:(ncol(data) - 1)) {
  nk[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_NK.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = nk, ncol = 3, top = textGrob("Natural Killers", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

```

### Melanoma_GSE93722

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

GSE93722_facs <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Facs_Melanoma_GSE93722.txt")
GSE93722_deconv <- read.delim("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/deconvolutions_pipeline/all_deconvolutions/all_deconvolutions_Melanoma_GSE93722.txt")

colnames(GSE93722_deconv) <- colnames(GSE93722_deconv) %>% str_replace_all("(CBSX__Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix_)|(CBSX_Fig2ab.NSCLC_PBMCs_scRNAseq_sigmatrix)", "CBSX.NSCLC")
colnames(GSE93722_deconv) <- colnames(GSE93722_deconv) %>% str_replace_all("(CBSX__scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b_)|(CBSX_scRNA.Seq_melanoma_Tirosh_sigmatrix_SuppFig_3.b)", "CBSX.melanoma")
colnames(GSE93722_deconv) <- colnames(GSE93722_deconv) %>% str_replace_all("(CBSX__sigmatrix_HNSCC_Fig2cd_)|(CBSX_sigmatrix_HNSCC_Fig2cd)", "CBSX.HNSCC")
colnames(GSE93722_deconv) <- colnames(GSE93722_deconv) %>% str_replace_all("CBSX__LM22_", "CBSX.LM22")

GSE93722_deconv <- GSE93722_deconv %>% select(-colnames(GSE93722_deconv)[grepl("XCELL", colnames(GSE93722_deconv))])
GSE93722_deconv <- GSE93722_deconv %>% select(-colnames(GSE93722_deconv)[grepl("MCP", colnames(GSE93722_deconv))])

# CD4
cd4 <- list()
data <- GSE93722_deconv %>% select(colnames(GSE93722_deconv)[grepl("CD4", colnames(GSE93722_deconv))])
# Regroup and remove different types of CD4 from LM22 signature
data$Epidish_CBSX_LM22_CD4 = data$Epidish_CBSX_LM22_T.cells.CD4.memory.activated + data$Epidish_CBSX_LM22_T.cells.CD4.memory.resting + data$Epidish_CBSX_LM22_T.cells.CD4.naive
data$DeconRNASeq_CBSX_LM22_CD4 = data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting + data$DeconRNASeq_CBSX_LM22_T.cells.CD4.naive
data$CBSX.LM22_CD4 = data$CBSX.LM22_T_cells_CD4_memory_activated + data$CBSX.LM22_T_cells_CD4_memory_resting + data$CBSX.LM22_T_cells_CD4_naive
data <- data %>% select(-c("CBSX.LM22_T_cells_CD4_memory_activated", "CBSX.LM22_T_cells_CD4_memory_resting", "CBSX.LM22_T_cells_CD4_naive", 
                           "Epidish_CBSX_LM22_T.cells.CD4.memory.activated" ,"Epidish_CBSX_LM22_T.cells.CD4.memory.resting" ,"Epidish_CBSX_LM22_T.cells.CD4.naive", 
                           "DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.activated" ,"DeconRNASeq_CBSX_LM22_T.cells.CD4.memory.resting" , "DeconRNASeq_CBSX_LM22_T.cells.CD4.naive"))

data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE93722_facs$CD4_Tcells)
for (i in 1:(ncol(data) - 1)) {
  cd4[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD4.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd4, ncol = 3, top = textGrob("CD4", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# CD8
cd8 <- list()
data <- GSE93722_deconv %>% select(colnames(GSE93722_deconv)[grepl("CD8", colnames(GSE93722_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE93722_facs$CD8_Tcells)
for (i in 1:(ncol(data) - 1)) {
  cd8[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(CD8.*)|(_T_cell.*)|(_T.cell.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = cd8, ncol = 3, top = textGrob("CD8", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# B
b <- list()
data <- GSE93722_deconv %>% select(colnames(GSE93722_deconv)[grepl("(B_cell)|(B.cell)|(B_lin)", colnames(GSE93722_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE93722_facs$Bcells)
for (i in 1:(ncol(data) - 1)) {
  b[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_(B_cell.*)|(_B.cell.*)|(_B_lin.*)")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = b, ncol = 3, top = textGrob("B cells", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

# NK
nk <- list()
data <- GSE93722_deconv %>% select(colnames(GSE93722_deconv)[grepl("NK", colnames(GSE93722_deconv))])
data <- data %>% select(order(colnames(.)))
data <- cbind(data, GSE93722_facs$NKcells)
for (i in 1:(ncol(data) - 1)) {
  nk[[i]] <- local({
    i <- i
    x = ggplot(data, aes(x = data[, ncol(data)], y = data[, i])) +
      geom_point() +
      stat_smooth(method = lm, size = 0.2, color = "black", se = FALSE, fullrange = T) +
      scale_x_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.0, 0.2)) +
      scale_y_continuous(limits = c(0, max(data[, i])), breaks = seq(0.0, 1.2, 0.2)) +
      theme(axis.title.x= element_text(size = 12, face="bold"), 
          axis.title.y= element_text(size = 12, face="bold") )  + 
      ylab("") + xlab("") + ggtitle(str_remove_all(names(data)[i], "_NK.*")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(color = "black", size = 14, face = "bold.italic", hjust = 0.5, vjust = 1.5)) +
      annotate("text", x = -Inf, y = Inf, hjust=-0.5, vjust=1, label = paste("italic(r)==", round(cor.test(data[, i], data[, ncol(data)])$est, 2), "~','", "~italic(P)<", "2.2e-16"), parse = TRUE, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=-1, vjust=2, label = paste("RMSE =", round(sqrt(mean((data[, i] - data[, ncol(data)])^2, na.rm = T)), 2)), parse = F, size = 6) +
      annotate("text", x = -Inf, y = Inf, hjust=2, vjust=3, label = paste("n = ", nrow(data)), size = 6)
  })
}
grid.arrange(grobs = nk, ncol = 3, top = textGrob("Natural Killers", gp=gpar(fontsize=20,font=3,face="bold")), left = textGrob("Deconvolution", gp=gpar(fontsize=14, face="bold"), rot=90), bottom= textGrob("FACS", gp=gpar(fontsize=14, face="bold")))

```