### 2021-12-14 Final_methylation_RNAseq_signatures.Rmd
Add files via upload

  ---
title: "R Notebook"
output: html_notebook
---
### To do different gene expression by limma package

```{r}

  DGEs <- function(data, Phenotype, lFC, FDR, method){
      require(limma)
      require(dplyr)
      #require(dendextend)

      message("[===========================]")
      message("[<<<<<<< DGEs START >>>>>>>>>]")
      message("[<<<< Pairwise analysis >>>>>]")
      message("------------------------------")

      # This function creates the pairs for the pairwise matrices
      design.pairs <- function(levels) {
          n <- length(levels)
          design <- matrix(0,n,choose(n,2))
          rownames(design) <- levels
          colnames(design) <- 1:choose(n,2)
          k <- 0
          for (i in 1:(n - 1))
              for (j in (i + 1):n) {
                  k <- k + 1
                  design[i,k] <- 1
                  design[j,k] <- -1
                  colnames(design)[k] <- paste(levels[i], "-", levels[j],sep = "")
              }
          design
      }

      # This function creates the pairs for the pairwise matrices

      design <- model.matrix(~0 + Phenotype)
      contr.matrix <- design.pairs(levels(factor(Phenotype)))
      colnames(design) <- rownames(contr.matrix)

      # Removing heteroscedascity from data
      v <- voom(log2(data+1), design, plot = F)

    if (method == "Blood") {  
      # Fitting linear models for comparisons of interest
      Fit <- lmFit(v, design) %>%
          contrasts.fit(., contr.matrix) %>%
          eBayes(.)

    } else if (method == "Cancer") {

      Fit <- lmFit(v, design) %>%
          contrasts.fit(., contr.matrix) %>%
          eBayes(.) %>% treat(., lfc = lFC)

    }

      FitList <- list()
      for (i in 1:ncol(contr.matrix)) {
          FitList[[i]] <- topTreat(Fit[which(decideTests(Fit)[,i] != 0),], coef = i, adjust.method = "BH", number = nrow(v$E)) %>% mutate(ID = rownames(.)) %>% filter(adj.P.Val < FDR)

      }

      names(FitList) <- colnames(contr.matrix)
      message("thank you for you waiting")
      return(FitList)

  }

  Signfeature <- function(data_TPM, Phenotype, FitList, FCcuoff ,MaxDMRs) {
      require(dplyr)
      require(matrixStats)
      require(limma)

      message("[===========================]")
      message("[<<<< Signature START >>>>>]")
      message("-----------------------------")

      # design <- model.matrix(~0 + Phenotype)
      # v_expr <- voom(data, design, plot = FALSE)

      # using positive FC (=over-expressed in cell type of interest)
      # FitList is DGEs which calculate by DGEs function

      A1 <- list()
      for (i in 1:length(FitList)) {

          A1[[i]] <- filter(FitList[[i]],  abs(logFC) > FCcuoff) %>% arrange(., desc(logFC)) #%>% top_n(ceiling(n()*0.5), wt = logFC)
          if (nrow(A1[[i]]) > MaxDMRs) {
              A1[[i]] <- A1[[i]][1:MaxDMRs,]
          }
      }

      sig <- lapply(A1, function(x) dplyr::select(x,ID))
      sig <- do.call(rbind, sig)
      sig <- filter(sig, !duplicated(ID))
      data1 <- data_TPM[rownames(data_TPM) %in% sig$ID,]

      # Print number of selected probes (signature)
      nrow(data1)

      result <- getMedVal(data1, Phenotype)
      message("[===========================]")
      message("[<<<< Signature END >>>>>]")
      message("-----------------------------")
      return(result)
  }

getMedVal <- function(data, Phenotype){
  library(matrixStats)

  Trans <- data.frame(t(data))
  Mt.Split <- split(Trans, Phenotype)
  Mt.Split <- lapply(Mt.Split, function(x) colMedians(data.matrix(x)))
  Mt.Split <- do.call(cbind, Mt.Split)
  rownames(Mt.Split) <- rownames(data)
  return(Mt.Split)
}

### To choose suitable threshold of LogFC (save as PDF file)

DGEs.QC <- function(FitList, FDR) {

    require(ggpubr)

    p <- list()

    for (i in 1:length(FitList)) {
        p[[i]] <- ggplot(FitList[[i]][FitList[[i]]$adj.P.Val < FDR,], aes(x = logFC)) + geom_density(colour = "red") + labs(title = names(FitList)[i], x = "logFC", y = "Density") + theme(plot.title = element_text(hjust = 0.5))
    }
    return(do.call(ggarrange, p))
}

getExpMatrix <- function(geneList, expression){

    df_out <- expression[geneList, ]
    return(df_out)
}

```

### BPRNA signature

```{r}

blueprint_DGEs_blood <- DGEs(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,7178:7216], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[7178:7216], FDR = 0.05, method = "Blood")

Mediant_blueprint_blood <- getMedVal(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,7178:7216], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[7178:7216])

BPRNA <- Signfeature(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(7178:7216)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[c(7178:7216)], FitList = blueprint_DGEs_blood,  FCcuoff = 2.5, MaxDMRs = 200)

```

### BPRNACan signature

```{r}

TCGA_normal_cancer_WB_DGEs <- DGEs(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(1:7164)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[1:7164], lFC = 1.5, FDR = 0.05, method = "Cancer")

blueprint_DGEs <- DGEs(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,7165:7216], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[7165:7216], lFC = 2, FDR = 0.05, method = "Cancer")

TCGA_cancer_blueprint_DGEs <- DGEs(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(684:6778, 7165:7216)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[c(684:6778, 7165:7216)], lFC = 2, FDR = 0.05, method = "Cancer")

Mediant_TCGA_normal_cancer_WB <- getMedVal(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(1:7164)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[1:7164])

Mediant_TCGA_cancer_blueprint <- getMedVal(data = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(684:6778, 7165:7216)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[c(684:6778, 7165:7216)])

Tumor.sig <- Signfeature(data_TPM = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(1:7164)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[1:7164], FitList = TCGA_normal_cancer_WB_DGEs, FCcuoff = 3, MaxDMRs = 100)

TIL.sig <- Signfeature(data_TPM = TCGA_normal_cancer_WB_blueprint_TPM_normalization[,c(684:6778, 7165:7216)], Phenotype = TCGA_normal_cancer_WB_blueprint_phenotype[c(684:6778, 7165:7216)], FitList = blueprint_DGEs, FCcuoff = 2, MaxDMRs = 150)

Immune_Tumor_CpGs <- unique(c(rownames(Tumor.sig), rownames(TIL.sig)))

BPRNACan <- getExpMatrix(geneList = Immune_Tumor_CpGs, expression = Mediant_TCGA_cancer_blueprint)

```