### 2021-12-14 Final_methylation_DNAm_signatures.Rmd
Add files via upload

---
title: "R Notebook"
output: html_notebook
---

### import bigwig data  ###

```{r}

# Function for importing bigWig data

FeatureBigWig <- function(x) {
	coverage <- c()
	for (i in 1:length(x)) {
    coverage <- c(coverage, list(data.frame(import(x[[i]]))))
  }
	coverage
}

# Merge the same samples's data

combine <- function(x){
    k <- c()
    for (i in seq(from = 1, to = length(x), by = 2)) {
        k <- c(k, list(data.frame(merge(x[[i]], x[[i + 1]], by.x = c('seqnames', 'start', 'end'), by.y = c('seqnames', 'start', 'end')))))
  }
    k
}

# Discards bases that have coverage below 10X

filter <- function(cg){
    flt <- c()
        for (i in 1:length(cg)) { 
        flt <- c(flt, list(data.frame(cg[[i]][cg[[i]]$score.y <= quantile(cg[[i]]$score.y, .999) & cg[[i]]$score.y >= 10,])))
        }
    flt
}

# Merge the all samples

final_merge <- function(path) {

    library(rtracklayer)

    files <- dir(path,  full.names = TRUE)

    Feature.cell <- FeatureBigWig(files)

    merge.cell <- combine(Feature.cell)

    filter.merge.cell <- filter(merge.cell)

    names.files <- c(names(filter.merge.cell[[1]])[1:3],unique(gsub('.CPG.*', '', dir(path))))

    dat = filter.merge.cell[[1]]

for (i in 2:length(filter.merge.cell)) {
    dat <- merge(dat, filter.merge.cell[[i]], by.x = c('seqnames', 'start', 'end'), by.y = c('seqnames', 'start', 'end'))
    }
    a <- dat[,1:3]
    for (j in seq(from = 6, to = ncol(dat), by = 6)) {
        a <- cbind(a,dat[, j])
        }
    colnames(a) <- names.files 
    a
}

```

# Generate signatures codes

```{r}

# Find different methylation positions (DMPs)

  DMPs <- function(data, Phenotype, method){
      require(limma)
      require(dplyr)

      message("[===========================]")
      message("[<<<< DMPs START >>>>>]")
      message("-----------------------------")

      design.pairs <- function(levels) {
          n <- length(levels)
          design <- matrix(0,n,choose(n,2))
          rownames(design) <- levels
          colnames(design) <- 1:choose(n,2)
          k <- 0
          for (i in 1:(n - 1))
              for (j in (i + 1):n) {
                  k <- k + 1
                  design[i,k] <- 1
                  design[j,k] <- -1
                  colnames(design)[k] <- paste(levels[i], "-", levels[j],sep = "")
              }
          design
      }

          design <- model.matrix(~0 + Phenotype)
          contr.matrix <- design.pairs(levels(factor(Phenotype)))
          colnames(design) <- rownames(contr.matrix)

      Fit <- lmFit(data, design) %>%
          contrasts.fit(., contrasts = contr.matrix) %>%
          eBayes(.)

      if (method == "TM") {

        FitList <- list()
        for (i in 1:ncol(contr.matrix)) {
            FitList[[i]] <- topTable(Fit[which(decideTests(Fit)[,i] != 0),], coef = i, adjust.method = "BH", number = nrow(data)) %>% mutate(ID = rownames(.))
            message(paste0(i, " done"))
        }

        } else if (method == "TIL") {
            FitList <- list()
            for (i in 1:ncol(contr.matrix)) {
               FitList[[i]] <- topTable(Fit, coef = i, adjust.method = "BH", number = nrow(data)) %>% mutate(ID = rownames(.)) 
              message(paste0(i, " done"))

        }
      }
      names(FitList) <- colnames(contr.matrix)

      message("[<<<<< DMPs END >>>>>>]")
      message("[===========================]")
      return(FitList)
  }

  # Function to get median value of each gene from an input dataframe

  getMedVal <- function(data, Phenotype){
    library(matrixStats)
    Trans <- data.frame(t(data))
    Mt.Split <- split(Trans, Phenotype)
    Mt.Split <- lapply(Mt.Split, function(x) colMedians(data.matrix(x)))
    Mt.Split <- do.call(cbind, Mt.Split)
    rownames(Mt.Split) <- rownames(data)
    return(Mt.Split)
  }

  Signfeature <- function(data, Phenotype, FitList, FDR, FCcuoff, MaxDMRs, method) {
    require(dplyr)
    require(matrixStats)
    require(limma)

      message("[===========================]")
      message("[<<<< Signature START >>>>>]")
      message("-----------------------------")

  # Filter by thresholds
  if (method == "TM") {
       A1 <- list()
      for (i in 1:length(FitList)) {
           A1[[i]] <- filter(FitList[[i]], adj.P.Val < FDR & abs(logFC) > FCcuoff) #%>% arrange(., desc(logFC))
          if (nrow(A1[[i]]) > MaxDMRs) { 
            A1[[i]] <- A1[[i]][1:MaxDMRs,]
            }
      }
  } else if (method == "TIL") {
      A1 <- list()
      for (i in 1:length(FitList)) {
          A1[[i]] <- filter(FitList[[i]], adj.P.Val < FDR & abs(logFC) > FCcuoff) #%>% arrange(., desc(logFC))
          if (nrow(A1[[i]]) > MaxDMRs) { 
            A1[[i]] <- A1[[i]][1:MaxDMRs,]
            }
      }
  }

    sig <- lapply(A1, function(x) dplyr::select(x,ID))
    sig <- do.call(rbind, sig)
    sig <- filter(sig, !duplicated(ID))
    data1 <- data[rownames(data) %in% sig$ID,]

  # Print number of selected probes (signature)
   nrow(data1)

    result <- getMedVal(data1, Phenotype)
      message("[===========================]")
      message("[<<<< Signature END >>>>>]")
      message("-----------------------------")
     return(result)
  }

  # function to extract expression values of a list of CpGs

  getExpMatrix <- function(geneList, expression){

      df_out <- expression[geneList, ]
      return(df_out)
  }

DGEs.QC <- function(FitList, FDR) {

    require(ggpubr)

    p <- list()

    for (i in 1:length(FitList)) {
        p[[i]] <- ggplot(FitList[[i]][FitList[[i]]$adj.P.Val < FDR,], aes(x = logFC)) + geom_density(colour = "red") + labs(title = names(FitList)[i], x = "logFC", y = "Density") + theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = seq(-1, 1, 0.2)) + theme(axis.text.x = element_text(size = 10, color = "Black"), axis.text.y = element_text(size = 10, color = "Black"))
    }
    return(do.call(ggarrange, p))
}

```

## BPmet signature ##

```{r}

FitList_IC_blood <- DMPs(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[, c(56:80, 93:105)], Phenotype = blueprint_pheno[c(1:25, 38:50)], method = "TIL")

Mediant_IC_blood <- getMedVal(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,c(56:80, 93:105)], Phenotype = blueprint_pheno[-c(26:37,51:52)])

BPmet <- Signfeature(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,c(56:80, 93:105)], Phenotype = blueprint_pheno[-c(26:37,51:52)], FitList = FitList_IC_blood, FDR = 0.05, FCcuoff = 0.2, MaxDMRs = 100, method = "TIL")

```

## BPmetCan signature ##

```{r}

  group <- c(rep("normal", 17), rep("cancer", 34), rep("wb", 4))

  Pheno2 <- c(as.character(group[18:51]), as.character(blueprint_pheno))

  FitList_TM <- DMPs(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,-c(56:107)], Phenotype = group, method = "TM")

  # FitList_IC <- DMPs(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,-c(1:55,81:92,106:107)], Phenotype = blueprint_pheno[-c(26:37,51:52)], method = "TIL")

  FitList_TIC <- DMPs(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,c(18:51,56:107)], Phenotype = Pheno2, method = "TIL")

Mediant_T <- getMedVal(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,-c(56:107)], Phenotype = group)

# Mediant_IC <- getMedVal(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,-c(1:55,81:92,106:107)], Phenotype = blueprint_pheno[-c(26:37,51:52)])

Mediant_TIC <- getMedVal(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,c(18:51, 56:107)], Phenotype = Pheno2)

Tumor.sig <- Signfeature(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,-c(56:107)], Phenotype = group, FitList = FitList_TM, FDR = 0.05, FCcuoff = 0.5, MaxDMRs = 100, method = "TM")

TIC.sig <- Signfeature(data = Normal_cancer_WB_blueprint_WGBS_genomic_numeric[,c(18:51,56:107)], Phenotype = Pheno2, FitList = FitList_TIC, FDR = 10e-6, FCcuoff = 0.3, MaxDMRs = 300, method = "TIL")

Immune_Tumor_CpGs <- unique(c(rownames(Tumor.sig), rownames(TIC.sig)))

BPmetCan <- getExpMatrix(geneList = Immune_Tumor_CpGs, expression = Mediant_TIC)

```