### 2022-01-26 Figures/Figure4.Rmd
Fix style in Figure4.Rmd

---
title: "Figure 4"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 15, fig.height = 17, echo = FALSE, results = "hide", warning = FALSE, message = FALSE)
```

```{r}
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(ggpubr)
library(grid)
library(gridExtra)
library(magrittr)
```

```{r}
plot_figure <- function(longer_merge, title) {
  longer_merge %<>%
    mutate(name = str_replace(name, "Avg_", "Average ")) %>%
    mutate(name = str_replace(name, "Met_", "Methylation "))
  ggplot(longer_merge, aes(sorted_type, fct_rev(name))) +
    geom_tile(aes(fill = cor), colour = "grey", size = 1) +
    scale_fill_gradient2(low = "#5C5DAF", mid = "white", high = "#EA2E2D") +
    geom_text(aes(label = text), col = "black", size = 8) +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.y = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 14),
      axis.text.x = element_text(
        angle = 90, hjust = 0.5, vjust = 0.5,
        size = 16, colour = "black", face = "bold"
      ),
      axis.text.y = element_text(size = 16, colour = "black", face = "bold")
    ) +
    labs(fill = str_c("Correlation")) +
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
}

prepare_table <- function(correlations, pvalues) {
  correlations %<>% select(...1, order(.[9, -1], decreasing = T))
  pvalues %<>% dplyr::select(colnames(correlations))

  longer_correlations <- correlations %>%
    pivot_longer(-1) %>%
    arrange(name)
  longer_pvalues <- pvalues %>% pivot_longer(-1)

  merged_all <- bind_cols(longer_correlations, longer_pvalues[, 3]) %>% dplyr::rename(cor = value...3, pvalue = value...4)
  merged_all %>%
    mutate(text = case_when(
      is.na(pvalue) & is.na(cor) ~ str_c("NA"),
      is.na(pvalue) & (cor == 0) ~ str_c(cor, "\n"),
      pvalue >= 0.05 ~ str_c(cor),
      pvalue >= 0.01 & pvalue < 0.05 ~ str_c(cor, "\n*"),
      pvalue >= 0.001 & pvalue < 0.01 ~ str_c(cor, "\n**"),
      pvalue < 0.001 ~ str_c(cor, "\n***")
    )) %>%
    mutate(sorted_type = factor(...1, levels = unique(...1)))
}
```

# Table of correlation

### Lymphocytes
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
lymph_correlations <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/lymph.csv")
lymph_pvalues <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/lymph_p.csv")

lymph_figure <- plot_figure(prepare_table(lymph_correlations, lymph_pvalues), "Lymphocytes")

# ggsave("lymph_figure.pdf", dpi = "print", width = 210, height = 297, units = "mm", scale = 2)
```

### Macrophages
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
macro_correlations <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/macro.csv")
macro_pvalues <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/macro_p.csv")

macro_figure <- plot_figure(prepare_table(macro_correlations, macro_pvalues), "Macrophages")

```

### Neutrophils
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
neutro_correlations <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/neutro.csv")
neutro_pvalues <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/all_tcga_corr/neutro_p.csv")

neutro_figure <- plot_figure(prepare_table(neutro_correlations, neutro_pvalues), "Neutrophils")

```
```{r}
ggsave("all_figure4_heatmaps.pdf", arrangeGrob(grobs = list(lymph_figure, macro_figure, neutro_figure)), dpi = "print", width = 210, height = 297, units = "mm", scale = 2)
```