### 2022-02-11 Figures/Figure2.Rmd
wrap tumour purity scatterplot + fix ridgeplot text adjust + loop through cancers

---
title: "Figure 2 and figure 2 supplementary material"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.width = 15, fig.height = 17, echo = FALSE, results = "hide", warning = FALSE, message = FALSE)
```

# Imports

```{r}
library(tidyverse)
library(Hmisc)
library(ggridges)
```

# Functions

```{r}
# Does the typical deconvolution scatterplot comparing true proportions to certain deconvolution estimates. Called inside scatterplot_wrapper
scatterplot_truevsestimate <- function(merged_data, x, y, title, color){
    merged_data %>% dplyr::select(all_of(x),all_of(y)) %>%
    na.omit() -> to_plot

    to_plot.cor <- rcorr(data.matrix(to_plot[c(x,y)]), type = "pearson")

  if (to_plot.cor$P[x,y] == 0){
    p.val = "< 2.2e-16"
    } else {
      p.val = paste("=", signif(to_plot.cor$P[x,y],2))
    }
    to_plot.rmse = sqrt(colMeans((to_plot[,y] - to_plot[,x])^2, na.rm = T))

  stats <- paste("Pearson: R = ", round(to_plot.cor$r[x,y], 2),
               ", pval ", p.val, "\n",
               "RMSE = ", round(to_plot.rmse, 2), sep = "")

  ## plot
  your_plot <- ggplot(to_plot) +
    aes_string(x = x, y = y) +
    geom_point(color = color) +
    labs(title = title, subtitle = stats, x = "", y = "") +
    geom_smooth(method=lm) +
    theme_classic() +
    theme(
      plot.title=element_text(family='', face='bold', colour='black', size=25, margin=margin(t=50,b=-40)),
      plot.subtitle=element_text(family='', face='italic', colour='black', size=18, margin=margin(t=50,b=-40)),
      text = element_text(size=15)
  )

  return(your_plot)
}

# This function returns the Lymphocytes, Macrophages and Neutrophils summarized
# of a given method/signature.
# Works inside of ridgeplot_parse. it takes:
# -  'deconv_data': contains the deconvolution data to index and summarize. 
#
# -  'signature dictionary': contains the exact names of the method
#                           (if it works with one signature) or the signature
#                           (if it works with multiple methods). 
#
# - iterator: string to subscript the signatures_dict with either signature
#             or methods names depending where is it called
#
# - prefix: string to leave just celltype names when substracted to the columns
#             of deconv_data of the specified method (or method + signature).
#
# - celltype: name of the celltype to sumarize "Lymphocites", "Macrophages" or "Neutrophils"

ridgeplot_parse <- function(deconv_data, signatures_dictionary){

  results <- tibble(Samples = deconv_data$Samples)

  names_table <- deconv_data %>%
    dplyr::select(-c(Samples)) %>%
    names() %>%
    str_split_fixed("_", 2) %>%
    as_tibble(.name_repair = "unique")

  for (method in unique(names_table[,1,drop=T])){

    if (method %in% c("Epidish",
                      "DeconRNASeq",
                      "CBSX")){

      for (signature in names(signatures_dictionary)){ # iterates through all the signatures (the ones not present for the method will yield all 0)
        to_match <- paste(method, signature, sep = "_")

        for (celltype in names(signatures_dictionary[[signature]])){
          if (length(signatures_dictionary[[signature]][[celltype]]) > 0){
            name_of_summary <- paste(to_match, celltype, sep = "_")
            columns_to_summarize <- paste(to_match, signatures_dictionary[[signature]][[celltype]], sep = ".")
            # print(name_of_summary)
            # deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% colnames() %>% print()
            # cat("-------------------------\n")
            deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
          }
        }
  }
    } else{

    to_match <- method

    for (celltype in names(signatures_dictionary[[method]])){
      if (length(signatures_dictionary[[method]][[celltype]]) > 0){
          name_of_summary <- paste(to_match, celltype, sep = "_")
          columns_to_summarize <- paste(to_match, signatures_dictionary[[method]][[celltype]], sep = ".")

          deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
      }
    }
  }
  }
  results %>% dplyr::select(-Samples) %>% dplyr::select(which(!colSums(., na.rm=TRUE) %in% 0)) %>% bind_cols(Samples = results$Samples,.)
}

plot_ridgeplot <- function(data){
 data %>% dplyr::select(-c(ParticipantBarcode, Study)) %>% data.matrix() %>% rcorr() -> cor.table
  final_plot <- list()
  for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    substring_remove <- c(paste("_",ct, sep=""),
      "_Tirosh_sigmatrix_SuppFig_3-b",
      "_scRNA-Seq",
      "Fig2ab-",
      "_PBMCs_scRNAseq_sigmatrix",
      "sigmatrix_",
      "_Fig2cd")

    # filter correlations
    deconv_ct_names <- data %>% dplyr::select(matches(ct), -ct) %>% names()
    cor.table[["r"]][ct,deconv_ct_names] %>% sort() -> cor.ct.sorted
    names(cor.ct.sorted) <- str_remove_all(names(cor.ct.sorted), paste(substring_remove, collapse = "|"))
    cor.ct.sorted["H&E"] = 1

    # ridgeplot perse
    cols <- c("H&E" = "red", "CBSX" = "yellow", "Epidish" = "blue", "Quantiseq" = "pink", "DeconRNASeq" = "green")

    ct_subset = data %>%
      dplyr::select(matches(ct)) %>%
      rename("H&E" = ct) %>%
      gather() %>%
      mutate(key = str_remove_all(key, paste(substring_remove, collapse = "|")),
             key=factor(key, levels=names(cor.ct.sorted)),
             method = as.factor(str_extract(key, "[^_]+")))
  adjust_x <- max(ct_subset$value)
  final_plot[[ct]] <- print(ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = adjust_x, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_x_continuous(limits = c(0, (adjust_x + adjust_x*0.2))) +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type"))
  }
  return(final_plot)
}

wrapper_ridgeplot <- function(deconv_data, true_estimates, cancer_abreviation){

  # build the data
  true_estimates %>% dplyr::filter(Study == cancer_abreviation) %>%  na.exclude() %>%
    arrange(factor(ParticipantBarcode, levels = ParticipantBarcode))  -> true_subset_

  Comm_samples <- Reduce(intersect, list(deconv_data$Samples,
                                         true_subset_$ParticipantBarcode))
  true_subset_ %>% 
    dplyr::filter(ParticipantBarcode %in% Comm_samples) %>% 
    arrange(factor(ParticipantBarcode, levels = Comm_samples)) -> true_subset

  deconv_data  %>%
    dplyr::filter(Samples %in% Comm_samples) %>% 
    arrange(factor(Samples, levels = Comm_samples)) -> deconv_subset

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)

  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()
  true_subset %>% bind_cols(deconv_subset_summary) %>% select(-c(Samples)) -> data_toplot

  # calculate correlations
 return(plot_ridgeplot(data_toplot))
}

wrapper_tpurity <- function(deconv_data, true_estimates, cancer_abreviation){
  true_estimates %>% dplyr::filter(Cancer_type == cancer_abreviation, CPE != "NaN") %>% 
    arrange(factor(Sample_ID, levels = Sample_ID))  -> true_subset_

  Comm_samples <- Reduce(intersect, list(deconv_data$Samples,
                                         true_subset_$Sample_ID))
  true_subset_ %>% 
    dplyr::filter(Sample_ID %in% Comm_samples) %>% 
    arrange(factor(Sample_ID, levels = Comm_samples)) -> true_subset

  deconv_data %>% 
    dplyr::filter(Samples %in% Comm_samples) %>% 
    arrange(factor(Samples, levels = Comm_samples)) -> deconv_subset

  # Scatterplot CCLE_TIL10
  identical(true_subset$Sample_ID, deconv_subset$Samples) %>% stopifnot()
   true_subset %>% bind_cols(deconv_subset) -> merged_subset

  final_plot = list()
  color = "black"

  title = "ESTIMATE vs CCLE TIL10 Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_CCLE_TIL10_Cancer.cells", title, color)

  # Scatterplot BPRNACan
  title = "ESTIMATE vs BPRNACan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_BPRNACan_cancer", title, color)

  return(final_plot)
}

summary_corrplots

```

# Data loading

```{r}
read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/julien/gemdecan/data/all_deconvolutions/all_deconvolutions_all_TCGA.txt") %>%
  mutate_at("Samples", str_remove_all, ".01A*") %>%
  mutate_at("Samples", str_replace_all, "-", ".") %>% 
  group_by(Samples) %>%
  summarise_all(mean) -> tcga_deconv

read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Aran_facs.csv") %>%
  mutate_at("Sample_ID", str_remove_all, ".01A*") %>%
  mutate_at("Sample_ID", str_replace_all, "-", ".") -> Aran_all_facs

read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv") %>%
  dplyr::select(-c(...1)) %>%
  mutate_at("ParticipantBarcode", str_remove_all, ".01A*") %>%
  mutate_at("ParticipantBarcode", str_replace_all, "-", ".") %>%
  distinct(ParticipantBarcode, .keep_all= TRUE) -> HE_all_TCGA # for the READ samples TCGA.AG.5 & TCGA.AG.3

```

# Figure
## A) Tumor Purity
```{r}
wrapper_tpurity(deconv_data = tcga_deconv,true_estimates = Aran_all_facs, cancer_abreviation = "LUAD")
```

## B) 3 Celltypes ridge plots
### Create aggregation dictionaries to compare with Lymphocites, Macrophages and Neutrophils FACs

```{r}
# Lists of celltypes of each signature
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
signature_dict <- list()
signature_dict[["BPRNACan"]] <- list(Lymphocytes = c("B", "CD4", "CD8", "NK"),
                    Macrophages = c("Mono", "M0", "M1", "M2"),
                    Neutrophils = c("Neu"))

signature_dict[["BPRNACanProMet"]] <- signature_dict[["BPRNACan"]]
signature_dict[["BPRNACan3DProMet"]] <- signature_dict[["BPRNACan"]]

signature_dict[["CBSX_Fig2ab-NSCLC_PBMCs_scRNAseq_sigmatrix"]] <- list(Lymphocytes = c("T.cells.CD8", "T.cells.CD4","NKT.cells" ,"B.cells", "NK.cells"),
                                                                         Macrophages = c("Monocytes"),
                                                                         Neutrophils = c())

signature_dict[["CBSX_LM22"]] <- list(Lymphocytes = c("B.cells.naive", "B.cells.memory",  "T.cells.CD8",
                                                        "T.cells.CD4.naive", "T.cells.CD4.memory.resting",
                                                        "T.cells.CD4.memory.activated", 
                                                        "T.cells.regulatory..Tregs.", "NK.cells.resting",
                                                        "NK.cells.activated"),
                                        Macrophages = c("Monocytes", "Macrophages.M0", "Macrophages.M1", "Macrophages.M2"),
                                        Neutrophils = c("Neutrophils")) # skipped , "Plasma.cells" "T.cells.follicular.helper", "T.cells.gamma.delta" in Lymphos, as well as "Dendritic.cells.resting", "Dendritic.cells.activated" for macros

signature_dict[["CBSX_scRNA-Seq_melanoma_Tirosh_sigmatrix_SuppFig_3-b"]] <- list(Lymphocytes = c("T.cells.CD8", "T.cells.CD4",
                                                                                                   "NK.cells", "B.cells"),
                                                                                   Macrophages = c("Macrophages"),
                                                                                   Neutrophils = c())

signature_dict[["CBSX_sigmatrix_HNSCC_Fig2cd"]] <- list(Lymphocytes = c("T.cells.CD8", "T.cells.CD4", "B.cells"),
                                                          Macrophages = c("Macrophages", "Dendritic.cells"),
                                                          Neutrophils = c())

signature_dict[["CCLE_TIL10"]] <- list(Lymphocytes = c("T.cells.CD8", "T.cells.CD4", "B.cells", "NK.cells", "Tregs"),
                    Macrophages = c("Macrophages.M1", "Macrophages.M2", "Monocytes", "Dendritic.cells"),
                    Neutrophils = c("Neutrophils"))

signature_dict[["Quantiseq"]] <- list(Lymphocytes = c("B.cell", "T.cell.CD4+.(non-regulatory)", "T.cell.CD8+", "T.cell.regulatory(Tregs)", "NK.cell"),
                    Macrophages = c("Macrophage.M1", "Macrophage.M2", "Myeloid.dendritic.cell", "Monocyte"),
                    Neutrophils = c("Neutrophil"))

# signature_dict[["MCP"]] <- list(Lymphocytes = c("T.cells", "CD8.T.cells", "Cytotoxic.lymphocytes","B.lineage" ,"NK.cells"),
#                     Macrophages = c("Myeloid.dendritic.cells", "Monocytic.lineage"),
#                     Neutrophils = c("Neutrophils")) # Based in enrichment so cant sum celltype estimates

# signature_dict[["XCELL"]] <- list(Lymphocytes = c(".B-cells", "CD4+.", "CD8+.", "CLP", "NK.", "Th1_cells", "Tgd_cells", "Th2_cells", "Tregs"),
#                     Macrophages = c("aDC", "cDC", "iDC", "pDC", "DC", "Macrophages.", "Monocytes"),
#                     Neutrophils = c("Neutrophils")) # Not in the paper in theory

```

### Data processing & ridge plot

```{r}

tcga_deconv %>%
  rename_at(vars(matches("^CBSX")), ~ paste0("CBSX_", .)) %>% 
  rename_at(vars(matches("__")), ~ str_replace_all(., "__", "_")) -> tcga_deconv_ridge

ridgeplots_by_cancer <- list()
tpurity_by_cancer <- list()
blank  <- ggplot() + theme_void()

for (cancer in unique(HE_all_TCGA$Study)){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank
}

ridges <- ggarrange(plotlist = ridgeplots_by_cancer[["LUAD"]], ncol = 3, common.legend = T, legend = "bottom")
scater <- ggarrange(plotlist = tpurity_by_cancer[["LUAD"]], ncol = 3, common.legend = T)
blank  <- ggplot() + theme_void()
fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
annotate_figure(fullplot, top = text_grob("LUAD", color = "black", size = 30))
pdf_name <- paste("test", "ridges", ".pdf", sep = "")
ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 3)
# for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
#    annotate_figure(print(ridgeplots_by_cancer[[cancer]][[ct]]),
#                              top = text_grob("Test", color = "black", size = 30),
#                              left = text_grob("Estimated fractions", color = "black", size = 20, rot = 90),
#                              bottom = text_grob("H&E estimates", color = "black", size = 20))
# }

#

```

## Summary Corrplot

```{r}

```

# Suplementary figure

```{r}

```